---
title: "ESM 204 HW 3"
author: "Patrick Pelegri-O'Day"
date: "5/2/2022"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(here)
library(tidyverse)
library(janitor)
```

### Starting facts

 - Consumers can be separated into two income groups: “high” and “low.” The data set provides price (in USD) and quantity (in kWh) estimates of demand per month for the two groups. Run linear regressions (with an intercept) to estimate the demand curves for “high” and “low” income consumers.
 - Initially, there is no tax on electricity consumption.
 - The current electricity price (without any taxes) is 0.10 USD per kWh.
 - The marginal cost of producing a kWh of electricity is linear and has a price-intercept of 0.

```{r}
# Read in energy data
energy_raw <- read_csv(here('data', 'HW3_data.csv'))

energy <- clean_names(energy_raw) %>% 
  select(-1) %>% 
  mutate(q_agg_kwh = q_high_kwh + q_low_kwh)
```


```{r}
# Run linear regression to estimate demand curves for high- and low-income consumers
low_lm <- lm(price_cents ~ q_low_kwh, data = energy)
high_lm <- lm(price_cents ~ q_high_kwh, data = energy)
```

### Question 1

One kWh of electricity emits 0.85 pounds of CO2. Assuming that the interim SCC correctly reflects the total social cost of one metric ton of CO2, what is the marginal externality cost per kwH of electricity?

- 1 kWh of electricity = 0.85 lb CO2
- 1 metric ton = 2204.6 lbs. Thus 1 metric ton of CO2 is emitted from 2204.6/0.85 = 2594 kWh of electricity
- SCC is $51/MT CO2. Thus, the SCC is set to 51 USD per 2594 kWh, or 1.97 cents per kWh.
- The marginal external cost of 1 kWh of electricity production considering the impacts of carbon emissions associated with that electricity production is **1.97 cents per kWh.** Note: there may be other negative external costs associated with non-carbon characteristics of electricity production such as land use. Those are not included in this external cost.

### Question 2

The aggregate monthly demand curve for electricity is found by horizontally stacking the demand curves for high income and low income consumers.

To find the supply curve, we want to know the quantity of kWh where aggregate demand equals 10 cents/kWh. This is 536,719.5 kWh. Since we know the origin of the supply curve is at y = 0 and the curve is linear, we can simply find the slope coefficient that connects the points (0 kWh, 0 cents/kWh) and (536,719.5 kWh, 10 cents/kWh).

**The supply and demand curves are graphed below.**

```{r}
# Functions

# Storing intercepts and slopes as objects
low_int <- low_lm$coefficients[1]
low_slp <- low_lm$coefficients[2]
high_int <- high_lm$coefficients[1]
high_slp <- high_lm$coefficients[2]

# Creating functions for both demand groups for graphing
d_low <- function(q) {low_slp*q + low_int}
d_high <- function(q) {high_slp*q + high_int}

# Creating an overall demand function for calculations
dem <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}

# Creating an aggregate demand function
dem_agg <- function(p){
  q <- dem(p, low_lm) + dem(p, high_lm)
  return(q)
}

# Creating a sequence to map the aggregate demand function for plotting
price = seq(0, 32, length.out = 100)
qagg <- map(price, dem_agg) %>% 
  unlist()

# Putting the sequence into a dataframe
df <- tibble(qagg = qagg, price = price)

# Storing the intercept and slope of the aggregate
d_agg_slp <- (0-low_int)/(dem_agg(0)-dem_agg(low_int))
d_agg_int <- 0 - d_agg_slp*dem_agg(0)
d_agg_plot <- function(q) {d_agg_slp*q + d_agg_int}

# Creating the MPC, MSC, MEC functions
mpc_slp <- (10/dem_agg(10)) # MPC slope
mpc <- function(q) {mpc_slp*q + 0} # MPC function
mpc_q <- function(p) {p/mpc_slp} # MPC in terms of Q
mec <- function(q) {1.97} # MEC function
msc <- function(q) {mpc(q) + mec(q)} # MSC function
msc_q <- function(p) {(p-1.97)/mpc_slp}

# Consumer surplus function
cs <- function(p, model){
  q <- dem(p, model)
  con_sur <- 0.5*(model$coefficients[[1]] - p)*q
  return(con_sur)
}

# Aggregate consumer surplus function
cs_agg <- function(p){
  cs <- cs(p,low_lm) + cs(p,high_lm)
  return(cs)
}

# Producer surplus function
ps <- function(p){
  q <- mpc_q(p)
  prod_sur <- 0.5*q*p
  return(prod_sur)
}
```

```{r}
# Plot

ggplot() +
  stat_function(color = "royalblue2", fun = d_low, size = 1, linetype = "twodash") +
  stat_function(color = "royalblue4", fun = d_high, size = 1, linetype = "twodash") +
  stat_function(color = "red3", fun = mpc, size = 0.75) +
  stat_function(color = "green4", fun = mec, size = 0.75) +
  geom_point(aes(mpc_q(10),10), size = 0.75, color = "blue") +
  geom_line(data = df, aes(x = qagg, y = price), color = "purple4", size = 1) +
  scale_x_continuous(name="Electricity (kWh)", limits=c(0,900000),expand = c(0, 0)) +
  scale_y_continuous(name="Price (cents)\n", limits=c(0,40),expand = c(0, 0)) +
  geom_text(aes(label = "Low Demand", x = 140000, y = 10), angle = -54, size = 3) +
  geom_text(aes(label = "High Demand", x = 370000, y = 14), angle = -33, size = 3) +
  geom_text(aes(label = "Aggregate Demand", x = 450000, y = 14.5), angle = -24, size = 3) +
  geom_text(aes(label = "MPC", x = 750000, y = 14.75), angle =11, size = 3) +
  geom_text(aes(label = "MEC", x = 850000, y = 3), size = 3) +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = c(0.95, 0.95), 
        legend.justification = c(1, 1),
        legend.spacing = unit(0, "cm"), 
        legend.margin = margin(0, 0, 0, 0, "cm"))
```


**Consumer surplus under status quo:** Total consumer surplus (using aggregate demand curve) at a price of 10 cents/kWh is **52,987.22 USD.** 

**Producer surplus under status quo:** Producer surplus at a price of 10 cents/kWh is **26,835.97 USD**

**Environmental cost under status quo:** 1.97 cents/kWh * 536,719.5 kWh = **10,573.37 USD**

### Question 3

Under the status quo, the **consumer surplus for low-income consumers is 8,112.43 USD** and the **surplus for high-income consumers is 44,874.79 USD.** These are calculated by finding the consumer surplus for the two consumer groups' respective demand curves at a price of 10 cents/kWh.

### Question 4

```{r}
# Demand function with tax for SCC at 51
dem_51 <- function(p, model){
  q <- (p - (model$coefficients[[1]]-1.97))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}

# Creating an aggregate demand function
dem_agg_51 <- function(p){
  q <- dem_51(p, low_lm) + dem_51(p, high_lm)
  return(q)
}
```


```{r eval = FALSE}
# Find the new equilibrium price at the intersection of dem_agg and msc
uniroot(function(p)
  dem_agg_51(p) - mpc_q(p),
        interval = c(0,20))
```

Applying a tax on consumers to address the interim SCC will shift the aggregate demand curve down by 1.97 cents/kWh. 
 
 - Setting the new aggregate demand curve equal to the supply curve, we find that the **new equilibrium price of electricity will be 9.32 cents/kWh.** This is the price received by producers. Consumers will pay 9.32 (producer price) + 1.97 (tax) = 11.29 cents/kWh.
 - Quantity of electricity: **500,213.2 kWh**, down from 536,719.5 kWh pre-tax
 - Welfare of high income consumers: **39,676.40 USD**, down from 44,874.79 USD pre-tax
 - Welfare of low income consumers: 6,622.60 USD surplus minus a cumulative external cost of 9,854.20 USD borne entirely by low-income consumers results in a **net loss of -1,741.95 USD**. If one treats  environmental damages before the tax as also entirely borne by low-income consumers, then pre-tax low-income consumer surplus pre-tax was -2,460.94 USD. Thus, low-income consumers are better off post-tax since their surplus is less negative.
 - Welfare of power suppliers (i.e. producers): **23,310.37 USD**
 - Total environmental damage: MEC times quantity produced = **9,854.20 USD**
 - Total tax revenue generated: size of the tax = MEC. So the total tax revenue generated will also be **9,854.20 USD**
 
### Question 5
 
```{r}
# calculate the cents per kWh of the SCC

scc_cents_kwh <- function(scc_cents) {
  cents_kwh <- round(scc_cents/2594, 2)
  return(cents_kwh)
}
```
 
```{r eval = FALSE}
############### 75
dem_75 <- function(p, model){
  q <- (p - (model$coefficients[[1]]-2.89))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}

dem_agg_75 <- function(p){
  q <- dem_75(p, low_lm) + dem_75(p, high_lm)
  return(q)
}

uniroot(function(p)
  dem_agg_75(p) - mpc_q(p),
        interval = c(0,20))

############## 100 
dem_100 <- function(p, model){
  q <- (p - (model$coefficients[[1]]-3.86))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
  
dem_agg_100 <- function(p){
  q <- dem_100(p, low_lm) + dem_100(p, high_lm)
  return(q)
}

uniroot(function(p)
  dem_agg_100(p) - mpc_q(p),
        interval = c(0,20))

############## 125
dem_125 <- function(p, model){
  q <- (p - (model$coefficients[[1]]-4.82))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
  
dem_agg_125 <- function(p){
  q <- dem_125(p, low_lm) + dem_125(p, high_lm)
  return(q)
}
  
uniroot(function(p)
  dem_agg_125(p) - mpc_q(p),
        interval = c(0,20))

############## 150
dem_150 <- function(p, model){
  q <- (p - (model$coefficients[[1]]-5.78))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
  
dem_agg_150 <- function(p){
  q <- dem_150(p, low_lm) + dem_150(p, high_lm)
  return(q)
}
  
uniroot(function(p)
  dem_agg_150(p) - mpc_q(p),
        interval = c(0,20))
```

 
 
```{r}
mec_51 <- function(q) {1.97} 
mec_75 <- function(q) {2.89} 
mec_100 <- function(q) {3.86}
mec_125 <- function(q) {4.82}
mec_150 <- function(q) {5.78}


```

 
```{r}

```

 
Dollars per kWh at each SCC. This the MEC and tax. This was found by converting USD/MTCO2 into USD/kWh.

 - 51: 1.97 cents/kWh
 - 75: 2.89 cents/kWh
 - 100: 3.86 cents/kWh
 - 125: 4.82 cents/kWh
 - 150: 5.78 cents/kWh
 
Equilibrium price (not including tax) and quantity at each dollar value of SCC. Aggregate demand functions were shifted downward by the tax calculated above. Then the intersection between the modified aggregate demand function and MPC curve was calculated to find equilibrium price and quantity.

 - 51 USD/MTCO2: 9.32 cents, 500213.2 kWh
 - 75 USD/MTCO2: 9.00 cents, 483233.5 kWh
 - 100 USD/MTCO2: 8.67 cents, 465121.9 kWh
 - 125 USD/MTCO2: 8.34 cents, 447293.3 kWh
 - 150 USD/MTCO2: 8.00 cents, 429747.6 kWh


```{r eval = FALSE}
for (i in c(9.32, 9, 8.67, 8.34, 8)){
  print(cs(i, low_lm))
}

for (i in c(9.32, 9, 8.67, 8.34, 8)){
  print(mpc_q(i))
     }
```


```{r eval = FALSE}
# calculate damage and total revenue
for (i in c(500222.5, 483047.5, 465335.8, 447624, 429375.6)){
  print
}

# calculate damage and total revenue
1.97 * 500222.5
2.89 * 483047.5
3.86 * 465335.8
4.82 * 447624
5.78 * 429375.6

for (i in c(9.32, 9, 8.67, 8.34, 8)){
  print(mpc_q(i))
     }
```

```{r eval = FALSE}
# price consumers pay with tax at different SCC levels
c(11.29, 11.89, 12.53, 13.16, 13.78)

# breakdown of consumption at different SCC levels
dem_51(11.29, high_lm)/(dem_51(11.29, high_lm) + dem_51(11.29, low_lm))
dem_100(12.53, high_lm)/(dem_100(12.53, high_lm) + dem_100(12.53, low_lm))
dem_125(13.16, high_lm)/(dem_125(13.16, high_lm) + dem_125(13.16, low_lm))
dem_150(13.78, high_lm)/(dem_150(13.78, high_lm) + dem_150(13.78, low_lm))
```


#### Calculating welfare

 - Welfare of high income consumers per estimated SCC was found by calculating the consumer surplus and adding the percent of electricity consumption by that group times the total tax revenue. 
  - Percent electricity consumption per group was found by taking the quantity of electricity a consumer group used at a given price and dividing by the quantity of electricity used by both consumer groups at that price.
 - Welfare of low income customers used the same method as welfare of high income consumers, followed by subtracting the total environmental cost (which is equal to the damage rate in cents/kWh times the total quantity of electricity consumed)
 - Welfare for producers was simply producer surplus with no adjustments needed.
 
```{r}
# scc_v <- c(51, 75, 100, 125, 150)
# 
# high_cons_v <- c(47,461.36, 48,675.21, 49,709.87, 50,832.16, 52,138.60)
# 

```

 
Welfare of high income consumers per estimated SCC

 - 51 USD/MTCO2: 3967640 (surplus) + 0.79 (pct consumption) x 985438.3 (tax revenue)  = **47,461.36 USD**
 - 75 USD/MTCO2: 3736755 + (0.81 x 1396007) = **48,675.21 USD**
 - 100 USD/MTCO2: 3498106  + (0.82 x 1796196) = **49,709.87 USD**
 - 125 USD/MTCO2: 3270876  + (0.84 x 2157548) = **50,832.16 USD**
 - 150 USD/MTCO2: 3054702  + (0.87 x 2481791) = **52,138.60 USD**
 - Note: The proportion of electricity consumption by high-income consumers rises as the effective price rises because high income consumers have a less elastic demand curve for electricity consumption.

 
Welfare of low income consumers per estimated SCC

 - 51 USD/MTCO2: 895855.2 (surplus) + 0.21 (pct consumption) * 985438.3 (tax revenue) - 985438.3 (env damages) = **1,173.59 USD**
 - 75 USD/MTCO2: 937124.7 + (0.19 x 1396007) - 1396007 = **-1,936.41 USD** (net loss)
 - 100 USD/MTCO2: 980657.1 + (0.18 x 1796196) - 1796196 = **-4,922.24 USD**
 - 125 USD/MTCO2: 1025178 + (0.16 x 2157548) - 2157548 = **-7,871.62 USD**
 - 150 USD/MTCO2: 1072081 + (0.13 x 2481791) - 2481791 = **-10,870.77 USD**
 
Welfare of electricity producers per estimated SCC

 - 51 USD/MTCO2: **23,310.37 USD**
 - 75 USD/MTCO2: **21,737.14 USD**
 - 100 USD/MTCO2: **20,172.31 USD**
 - 125 USD/MTCO2: **18,665.92 USD**
 - 150 USD/MTCO2: **17,175.02 USD**

### Question 6

I assume we use an SCC of 51 USD/MTCO2e and a tax of 1.97 cents/kWh. Adding solar for the high income group will halve the y-intercept of their demand curve after accounting for the tax. The y-intercept for the low income group's demand curve will only be affected by the tax.

 - The new quantity of electricity consumed is **313,444.2 kWh**
 - Total environmental externality is **6,174.85 USD**: the quantity of electricity consumed times the marginal external cost of 1.97 cents/kWh
 
 Assuming there is no solar available, the total size of tax that would lead to an equilibrium quantity of electricity consumption of 313,444.2 kWh is **12.05 cents/kWh** which would result in producers receiving 5.84 cents/kWh and consumers paying 17.89 cents/kWH

```{r}
# Demand function with solar and SCC at 51
dem_high_solar <- function(p){
  q <- (p - (high_lm$coefficients[[1]]-1.97)/2)/high_lm$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}

dem_low_solar <- function(p){
  q <- (p - (low_lm$coefficients[[1]]-1.97))/low_lm$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}

# Creating an aggregate demand function
dem_agg_solar <- function(p){
  q <- dem_high_solar(p) + dem_low_solar(p)
  return(q)
}
```



```{r eval = FALSE}
# Find the new equilibrium price at the intersection of dem_agg_solar and mpc
uniroot(function(p)
  dem_agg_solar(p) - mpc_q(p),
        interval = c(0,20))

# Find quantity of electricity consumed
mpc_q(5.84)
```

